---
title: "Are Younger Generations More Acceptant of Democratic Backsliding?"
author: "Olivier Bergeron-Boutin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
header-includes: 
  - \usepackage{graphicx}
  - \usepackage[doublespacing]{setspace}
  - \usepackage{lscape}
geometry: margin=1.5in
fontsize: 12pt
mainfont: Bembo Std
bibliography: youth.bib
biblio-style: apalike
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()#,
                     #kable.force.latex = TRUE
)
# kable.force.latex is to make knitr::kable caption appear in the sidenotes. 
```

The scholarly study of democratic erosion has long been reserved to those regimes that were considered most vulnerable -- the fledgling democracies whose institutions were weakly entrenched and whose peoples had not yet internalized the norms of popular government.^[The data and R scripts required to replicate the analyses presented in this paper can be found in a [github repository](https://github.com/aka09/youth_project)
.] In the past five years, however, the resilience of consolidated democracies has come under scrutiny [see e.g. @levitsky_how_2018;@kaufman_democratic_2019]. In particular, Foa and Mounk's [-@foa_danger_2016] claim that younger generations are less attached to democracy and more open to authoritarian alternatives has attracted much attention.  Amid concern about the future of democracy, this question is of critical importance and deserves a comprehensive empirical examination. Is Western democracy sustainable if its youngest citizens are tempted by the allure of authoritarianism? A long tradition in political science would argue that it cannot: a regime becomes consolidated only once it has become "the only game in town" [@linz_toward_1996]. The most robust empirical evidence produced to date also supports this assessment [@claassen_mood_2020]. Declining rates of support for democracy among youth would be particularly worrying if due to cohort effects, as opposed to life-cycle effects. Indeed, cohort dynamics imply that despite the effects of aging, today's younger cohorts would never reach the level of support that the current oldest generations express. In the long-run, through intergenerational replacement, population-level suppor for democracy would therefore decrease.

Recognizing the urgency of this research question, scholars have contributed to a burgeoning literature examining generational patterns of support for democracy in Europe and North America. 
Although it is noteworthy that most researchers strike a noticeably less pessimistic tone than Foa and Mounk [see e.g. @zilinsky_democratic_2019;@wuttke_have_2021], no single conclusion can claim universal support and the issue of youth support for democracy is still very much a contentious area of research. This research project aims to contribute to this literature by adopting a two-pronged approach. First, given the controversy surrounding the potential spuriousness of Foa and Mounk's findings, I use the traditional survey measures of support for democracy and seek to assess empirically the robustness of the link between age and attitudes toward democracy. My approach here is to use specification curve analysis, a recent method that aims to systematically examine how empirical findings change due to variation in modeling choices [@simonsohn_specification_2020]. The second half of the paper focuses on exploring alternative measures of support for democracy and the insights that they may yield with regard to the question of youth's commitment to democracy. In this section, I exploit original data from four surveys fielded in the United States and Canada in both 2019 and 2020. Embedded in these surveys were a series of conjoint experiments and traditional vignette experiments designed to test respondents' reactions to antidemocratic behavior on the part of political elites. By using a combination of more conventional measures and newly developed tools, my objective is to provide a new look into an important question: are younger generations more acceptant of democratic backsliding? Before turning to the empirical section, I begin with a short review of the key findings in this literature. 

# The literature so far

The ongoing controversy over youth's commitment to democracy can be traced back to Foa and Mounk's [-@foa_danger_2016;-@foa_signs_2017;-@foa_end_2017] series of influential articles in the \textit{Journal of Democracy}. Foa and Mounk build an argument against the so-called "consolidation paradigm" -- a scholarly approach that, they argue, is blind to the potential for deconsolidation among even the most advanced democracies. Foa and Mounk diagnose an instability to democratic regimes that have long been thought to be stable and unlikely to break down. Using data from the World Values Survey, they show troubling patterns: in a variety of advanced democracies, younger respondents are less likely to agree that it is important to live in a democracy, more likely to say that having a democratic political system is a bad way to govern, and more likely to endorse authoritarian alternatives to democracy, such as having a "strong leader who does not have to bother with Parliament and elections." Foa and Mounk's piece reads not only as a warning, but also as an indictment of the "consolidation paradigm" and its proponents, whom they accuse of having been asleep at the wheel. 

Foa and Mounk's original articles spurred a scholarly debate that is still ongoing. Their provocative claims have attracted so much attention that the \textit{Journal of Democracy} published a special online issue dedicated to a discussion of their findings. Erik Voeten [@voeten_are_2017] notably argued that Foa and Mounk's peculiar coding scheme, which collapsed answers that rated democracy as anything other than maximally important into a single category,  artificially inflated the cohort effects. Indeed, many among the younger cohorts rated the importance of democracy as 8 or 9, a still-strong level of support. Therefore, while the proportion of younger respondents who answered 10 was drastically lower, the differences in mean scores on the full 0 to 10 scale did not differ nearly as drastically. To make up my own mind about this, I analyzed data from Wave 6 of the European Values Survey, which contained an extensive module on attitudes toward different political regimes. 

Figure 1 shows the results: among 15 of the most developed countries in Europe, there is \textit{some} evidence that younger birth cohorts accord a lower importance to democracy.\footnote{Note that given the fairly high inter-country variation, the estimates for each country are differences between a given cohort's mean importance and the mean importance reported by individuals born before 1950 in that same country. For the two youngest birth cohorts, the differences that are statistically significantly different from the pre-1950s cohort are shown in red ($\alpha$ = 0.05).} For 9 of these 15 nations, the importance that the two youngest cohorts accord to democracy differs significantly from the pre-1950s cohort, which serves as the baseline. In no country do we find a significantly higher importance of democracy among the younger cohorts than among the oldest cohort. While this would seem to support Foa and Mounk's assertions, I wish to make two points in relation to Figure 1. First, there are 6 countries out of 15 for which there is no robust statistical evidence: Belgium, Italy, the Netherlands, Porgular, Spain, and Switzerland. Second, as Pippa Norris [-@norris_is_2017] notes, even when there \textit{are} statistically significant estimates, the differences are substantively small -- generally, far less than a full point on the 0 to 10 scale. In sum, this short descriptive exercise neither refutes, nor provides clear evidence in favor of, Foa and Mounk's claims. 

\begin{figure}
\caption{Stated importance of democracy by birth cohort}
\includegraphics{important~cohort+country.png}
\end{figure}

Additional critiques were made by Alexander and Welzel [-@alexander_myth_2017], who focus not on methodology but instead on the proper interpretation of Foa and Mounk's findings. Alexander and Welzel contend that younger cohorts' understanding of democracy may be fundamentally different. As such, an older respondent answering that democracy is "absolutely important" -- the maximum score of 10 -- may not hold the same attitudes as a younger respondent who provides the same answer. The younger cohorts' expressed support may be more meaningful because "the moral values on which people base their democratic support have turned dramatically more liberal over the generations" [@alexander_myth_2017,p.1]. Finally, in a more recent article, Wuttke et al. [-@wuttke_have_2021] use generalized additive models to tease apart the age, cohort, and period effects that may explain generational differences in support for democracy; they find no indication that the youngest cohorts are distinctive in this regard. Their findings imply that as today's young cohorts age, they will come to adopt attitudes that are similar to today's older cohorts. 

# Measuring support for democracy 

Despite somewhat contradictory results, the extant literature shares a reliance on survey questions that directly query respondents about their attitudes toward various political regimes. Such questions are widely used in cross-country, nationally representative surveys, which facilitates comparative analyses of attitudinal democratic backsliding. However, there is reason to doubt that they are effective at measuring the latent construct of support for democracy. I highlight three reasons why this is and provide suggestive evidence of their inadequacy.

First, although a handful of survey items are frequently used to measure support for democratic rule, \textit{which} of these popular items best measures the underlying concept remains unclear -- with important consequences on the conclusions that can be drawn. Notably, four survey items that are regularly used in studies of democratic backsliding are only weakly correlated (see Table 1).\footnote{Three items ask whether the following types of political systems are good ways of governing the respondent's country: "Having a strong leader who does not have to bother with parliament and elections", "Having the army rule the country", and "Having a democratic political system". The fourth item asks participants to rate the importance of living in a democratic country on a scale from 1 to 10.} Thus, there is only a tenuous link between different survey instruments that have been used interchangeably as dependent variables in the literature. Which of these four items -- if any -- best captures regime support is unclear, but what \textit{is} clear is that basic issues of measurement have not been given sufficient attention. In short, the most widely used measures of support for democracy are, at best, noisy signals. Therefore, relying on these items to conduct analyses of generational patterns may yield faulty inferences. I remain agnostic as to the particular nature of the biases that this may induce; the broader point is that the validity of survey measures must be demonstrated, rather than simply assumed. In particular, survey researchers that the vagaries of survey research should make us particularly wary to "a mismatch between the level of specificity in the attitudes that we would like people to have and in the attitudes they actually possess" [@berinsky_measuring_2017,p.311]. In the case of support for democracy, recent evidence suggests that "the inclusion of 'democracy' [in survey questions] focuses respondents' attention on regimes in abstract terms rather than forcing them to consider their orientations toward regimes in practice" [@kiewiet_de_jonge_should_2016,p.696].

```{r,results='asis'}
library(tidyverse)
library(knitr)
library(kableExtra)
# Cleaned EVS wave 5 data
load("cleaned_data/evs_w5.RData")
# Correlation matrix
bruh <- evs5 %>% 
  select(sys1_strongleader:sys3_democracy, dem_important) %>% 
  rename(`Strong leader` = sys1_strongleader,
         `Army rule` = sys2_army,
         `Democracy (GW)` = sys3_democracy,
         `Democracy (I)` = dem_important) %>% 
  cor(use = "complete.obs") %>% 
  round(3) 
  knitr::kable(bruh,
        caption = '\\textbf{Correlation matrix between 4 items measuring regime support.}',
        format = "latex",
        booktabs = TRUE) %>% 
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    add_footnote(c("Data from Wave 5 of the European Values Survey", "GW = Good way; I = Importance"))
```

Second, social desirability bias may produce a misleadingly optimistic picture of support for democracy. As social scientists have developed new methodological approaches in order to get around the problem of untruthfulness\footnote{Perhaps most notably, list experiments have been used as a statistical "truth serum", see @glynn_what_2013 and @lyall_explaining_2013}, research on support for democracy has mostly stuck with the same direct questions that may or may not elicit honest answers. Although problematic in its own right, this issue of sensitive survey questions may spell further trouble if the pressure to deliver the socially acceptable answer differs from generation to generation. If this is the case, the observed differences in support for democracy across cohorts may be an artifact of differential rates of social desirability bias.\footnote{There is, as far as I know, no robust evidence on this question in political science. Nonetheless, seriously considering the possibility is, I believe, crucial to the research agenda on youth and democracy. Again, I would argue that the validity of survey instruments must be demonstrated, not assumed. Using a list experiment to probe the presence of social desirability bias in questions about democracy may be a fruitful avenue of research.}

Third, expressing support for democracy in a survey is essentially costless. In fulfilling their democratic duties, citizens are never presented with a simple up-or-down vote on the continuation of democratic government. Instead, the respect of democratic norms is one of many considerations that must compete with, among others, ideology and partisanship. What matters most is not answers to abstract survey questions, but how citizens behave when they are asked to make trade-offs between policy preferences, partisan loyalty, and respect for democracy. If older generations hold intense preferences for certain policies or are less willing to cross party lines, the generational differences identified by the nominal measures may disappear. Real-world choices often involve a choice between an ideologically-proximate, but undemocratic candidate and an ideologically-distant who is respectful of democratic norms. Evaluating how citizens make these trade-offs is at the heart of the issue of democratic resilience; standard survey questions cannot provide any information about these trade-offs. As a result, these familiar items about regime types may identify as small-d democrats those respondents who defend democracy "in the abstract" but whose support crumbles due to the smallest inconvenience [@schedler_democrats_2007].

\begin{figure}
  \caption{Proportion of EVS Wave 5 respondents endorsing a strong  leader by stated importance of democracy}
  \includegraphics{strong~important+country.png}
\end{figure}

Figure 2 presents an analysis of this phenomenon using data from the European Values Survey. Respondents are divided into two groups: those who agree that a strong leader is a "very good" or "fairly good" way of governing their country, and those who do not.\footnote{As a reminder, the full formulation is: "a strong leader who does not have to bother with Parliament and elections."} The figure shows the proportion of EVS respondents in each country who fall in the former group, by the importance that they say they accord to democracy. It is apparent that a non-negligible proportion of citizens who should be the strongest defenders of democracy nonetheless endorse the strong leader solution. For instance, fully 25% Dutch respondents who answered 10 on the importance of democracy question agree that a strong leader would be a good way of governing their country. The Netherlands represents a bit of an extreme case\footnote{But note Portugal!}, but across most countries, about 1 in 6 EVS participants that rate democracy as maximally important nonetheless express support for this unambiguously authoritarian solution. As expected, the situation is even more troubling when looking at respondents who answered 9 on the importance of democracy question. We may of course expect these individuals to have \textit{some} reservations about democratic governance, but the picture that emerges from the EVS data indicates worrying openness to the strong leader solution that a "9" answer does not convey. For instance, in the United Kingdom, nearly 30% of respondents 9 on the importance of democracy question agree that a strong leader would be a good way of governing their country.\footnote{Note that the estimates have relatively wide confidence intervals given the extremely skewed distribution of answers on the importance of democracy question.} As such, it appears that the widely-used question about the importance of democracy fails to discriminate genuine small-d democrats from "democrats in the abstract." 

This conclusion is supported by original, nationally representative surveys in Canada and the United States fielded in May 2020. Figure 3 presents the same analysis as above, but this time restricting to respondents who answered 10 and breaking down the results by birth cohort. Strikingly, 44% of young Americans and 36% of young Canadians who say that democracy is absolutely important nonetheless endorse the strong leader solution. 

\begin{figure}
  \caption{Proportion of respondents who say that democracy is maximally important but who endorse a strong leader, by birth cohort}
  \includegraphics{strong~important+country+gen.png}
\end{figure}

In sum, direct questions asking about respondents' views of different political regimes are one way of measuring individual-level commitment to democracy. It is, to be sure, a convenient way, but also an imperfect one -- for the reasons stated above. Consequently, this research project uses a two-pronged approach, which I describe below. 

# Contribution: sensitivity analyses and experimental measures

## Specification curve analyses 

In the first part of this research project, I will remain within the confines of the traditional measures of support for democracy and probe the robustness of the empirical relationship between generational cohort and support for democracy. A concern that has been raised multiple times regarding Foa and Mounk's findings is that they are dependent on unusual coding schemes that inflate the generational differences in support for democracy. For instance, Erik Voeten [-@voeten_are_2017] argues against Foa and Mounk's decision to dichotomize the importance of democracy scale, with respondents who answered anything other than the maximum 10 being coded as a single group. These arbitrary decisions, which are almost inevitable in academic research and over which researchers often have much flexibility, can be exploited in order to produce results that satisfy the researcher. This is the familiar "researcher degrees of freedom" problem that has plagued the social sciences and contributed to the replicability crisis [@simmons_false-positive_2011]. In regression analysis, the arbitrary decisions that researchers must make include: which covariates to control for, the assumed functional form of the relationship between the independent variables and the outcome, which observations to exclude, which type of model to use, whether to include any interaction terms...As the number of decisions -- and therefore the number of potential specifications -- approaches infinity, the probability of finding a statistically significant relationship between the main independent variable and the outcome of interest converges toward 1. As such, one may worry that a regression-based observational study may "merely constitute a demonstration that is is possible to find a specification consistent with the author's prior expectations" [@ho_matching_2007,p.200].

In order to assess the robustness of the link between age and support for democracy, I propose to use specification curve analysis, a modern and increasingly popular technique to battle the problem of false positives in the social sciences.\footnote{I am not aware of any article in political science journals that use specification analysis, though such articles in psychology and economics journals seem to be relatively frequent.} The idea is to pre-specify a universe of "reasonable specifications" that contains all "valid and non-redundant" ways to model a relationship between some independent variable and some outcome variable [@simonsohn_specification_2020]. Once the data has been collected, the researcher can then systematically test all of these pre-specified models, plot the resulting point estimates from smallest to largest, and indicate the modeling choices that produced each point estimate. For instance, in the case that interests us, arbitrary decisions must be made with regard to defining/measuring generation cohorts. Different reasonable specifications could be presented: birth cohorts as decade of birth, birth cohorts using some conventional thresholds -- such as 1945-1959 for baby-boomers --, or age as a numeric variable -- with or without a squared term. All of these choices may -- or may not -- impact the substantive conclusions drawn on youth and democracy. The choice of outcome variable may also exert some influence: if we consider the four different items on support for democracy described earlier and the four operationalizations of birth cohorts, we already have 16 potential specifications.} One must then consider the relevant control variables, their functional form, the proper regression model...

Given the concerns highlighted earlier about the validity of the direct measures of support for democracy, a robust relationship demonstrated by specification curve analysis would not definitively "prove" the existence of a worrying trend regarding youth and democracy. Still, it would help assuage concerns that have been expressed about the quality of the empirical work presented by Foa and Mounk. For this first section of the paper, I use survey data collected in May 2020 in Canada and the United States (in collaboration with Profs. Dietlind Stolle and Elisabeth Gidengil).\footnote{An extension of the current analysis could draw on additional data sources, including similar original surveys that were fielded in 2019, as well as large-scale surveys such as the European Social Survey.} This survey, fielded in May 2020 in both countries during the coronavirus crisis, included 2,421 respondents in Canada and 2,727 respondents in the United States. We should thus have more than enough power to detect even quite subtle cohort differences. 

There are of course a variety of strategies that could be adopted here. I first examine the stated "importance of democracy" as a dependent variable. As a reminder, this variable ranges from 0 to 10, where 10 represents the maximum stated importance of democracy. On the right-hand side, I introduce age as the main explanatory factor of interest. Using age, instead of some categorical variable that classifies respondents into different birth cohorts, has the significant disadvantage of assuming a particular functional form. However, a notable advantage is that using age on the right-hand side will produce a single estimate of interest -- $\beta_{Age}$. In addition, I consider a variety of additional covariates: gender, authoritarianism, party identification, economic deprivation, democratic knowledge, COVID health worries, and COVID economic worries. Finally, because of repeated concerns that I have personally had regarding the attention that young respondents pay to surveys they participate in, I also divide the samples into 4 groups of equal size based on the time that each respondent spent answering the questionnaire.^[Note that the survey data already has multiple quality filters: the exclusion of "straightliners", the exclusion of respondents who answered the survey in fewer than 300 seconds, and the exclusion of respondents who failed an attention check.] If the age differences are only found among those who spent the least time answering the survey, we may suspect that inattention is driving the results.^[Of course, it is also possible that those who spend less time on the survey have genuine differences of opinion regarding democracy.]

The results of the specification curve analyses regarding the "importance of democracy" measures are presented separately for each country in Figures 4 and 5. Panel A of the figures present the different coefficients on age from 45 separate regression models. The y axis shows the coefficients on age and their associated 95% confidence intervals, while the x axis simply indexes the model number. The models are ordered so that that smallest coefficients on age are presented on the left, while the largest coefficients are presented on the right. With the stated importance of democracy as the dependent variable, a positive coefficient then indicates that as age increases, we would expect the stated importance of democracy to increase as well. The bottom panel shows the regression specifications that produced the different estimates, with vertical bars indicating that a particular covariate or subset of the data was used.^[All models are linear. Note that I deviate somewhat from the "reasonable and non-redundant" modeling approach described earlier. Since the goal of this research project is not expressly causal inference, I present a variety of what we may call "underspecified" models that include a single covariate.]

\begin{landscape}
\begin{figure}
\caption{Specification curve analysis for Canada, importance of democracy (2020 survey)}
\includegraphics[height=\paperwidth]{spec_imp_can.png}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Specification curve analysis for the United States, importance of democracy (2020 survey)}
\includegraphics[height=\paperwidth]{spec_imp_usa.png}
\end{figure}
\end{landscape}

Figures 4 and 5 indicate a robust relationship between age and stated importance of democracy. In Canada, every single coefficient is positive, and all but two are statistically significant. To get a sense of the substantive magnitude of the estimated effects, we can simply multiply what is a typical effect (0.02) by 40, the approximate age difference between the oldest cohort (pre-1950s) and the youngest cohort (post-1990s). The result, which is 0.80, suggests fairly large age patterns in stated importance of democracy. In an unexpected turn, it appears that respondents who rushed through the survey are *not* driving the results: the 7 largest estimates are found among the subset of respondents who fall in the 3rd quartile of survey duration. The conclusion is similar in the United States; if anything, the relationship appears more robust and of a greater substantive magnitude.

Figures 6 and 7 repeat the analysis, but instead using support for a strong leader as the dependent variable.^[The coding scheme is as follows: respondents who answered that a strong leader would be a "very good" way of governing their country were coded 1; "fairly good" answers were coded 0.66; "fairly bad" were coded 0.33; and "very bad" were coded 0.] Here, a negative coefficient indicates that as age increases, we expect less support for a strong leader (and conversly, as age decreases we expect more support for a strong leader). This is exactly what we find, albeit the relationship appears somewhat less robust than that examined earlier. 

The results of the specification curve analysis appear quite clear: Foa and Mounk's warnings find support in the data. But, as explained earlier, the validity of the very survey questions that are used by many scholars, as well as in this paper, is unclear. I therefore turn to the second part of this project, which examines alternative measures of regime support. 

\begin{landscape}
\begin{figure}
\caption{Specification curve analysis for Canada, strong leader (2020 survey)}
\includegraphics[height=\paperwidth]{spec_sl_can.png}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Specification curve analysis for the United States, strong leader (2020 survey)}
\includegraphics[height=\paperwidth]{spec_sl_usa.png}
\end{figure}
\end{landscape}



# Experimental evidence 

Given the limitations of the direct measures of support for democracy, this research project also aims to provide new evidence on youth and democracy by drawing on "behavioral" measures from survey experiments. The specification curve analysis presented in the previous section seems to confirm the robustness of the findings regarding youth and attitudes toward democracy. However, as my extensive critique of the so-called "nominal" measures of support for democracy foreshadows, I do not believe that the discussion should end there. I thus propose to measure support for democracy in a variety of alternative ways, all of which aim to sidestep the problems highlighted above. Notably, the focus here is on \textit{not} explicitly mentioning democracy and on emphasizing this tradeoff between policy and process preferences that I mentioned earlier. 

The first type of evidence that I present comes from conjoint experiments embedded in nationally representative surveys in Canada and the United States fielded in May 2020. The conjoint experiments presented pairs of fictitious candidates running for political office. Each candidate was randomly assigned 7 characteristics: age, sex, political experience, party label, position of coronavirus lockdowns, position on economic aid, and a position on the proper role of the executive branch of government. In a random half of the sample, this last feature was randomly assigned to represent positions on weakening legislative checks. The candidate was then described as either arguing that "A president should work with Congress even if it is obstructing his/her proposals to combat the pandemic" or that "A president should shut down Congress if it is obstructing his/her proposals to combat the pandemic." For the other random half sample, the options were focused on judicial checks: “A
president should comply with court decisions overturning his/her policies to combat a pandemic” or “A president should ignore court decisions overturning his/her policies to combat a pandemic." Respondents were then asked to make a choice between these two candidates based on their descriptions. 

The quantity of interest in this analysis is the average marginal component effect (AMCE), defined as the estimated change in probability that a profile (candidate) is selected caused by a change in some attribute (characteristic) from its reference level to some other level [@hainmueller_causal_2014]. The advantages of conjoint analysis are twofold. First, by presenting candidates that have an array of characteristics, respondents have plausible deniability when making choices that they perceive may be unpopular or socially undesirable. Second, by presenting candidates that vary on salient dimensions of vote choice -- such as policies related to the ongoing coronavirus crisis -- respondents must make difficult choices between an ideologically-congruent, but undemocratic candidate and an ideologically-distant, but democratic candidate. This setup, I argue, is better suited to simulating the conditions under which citizens make political decisions. 

\begin{figure}
  \caption{Average Marginal Component Effects by condition, generation, and country}
  \includegraphics{amces~gen.png}
\end{figure}

Results are presented in Figure 8, which shows average marginal components effects by democracy attribute, birth cohort, and country. In the case of this conjoint experiment, a negative AMCE attached to an antidemocratic attribute represents a punishment that voters hand out due to antidemocratic behavior on the part of candidates for office. In other words, assuming that one would like to observe a citizenry that ferociously defends demoratic government,we would like to observe \textit{negative} estimates -- and the larger, the better. In the United States, there is a (relatively) clear cohort pattern: as we move from the pre-1950s cohort to the post-1980s cohort, we see a roughly linear trend in the estimated AMCEs. The estimates for the oldest cohort are -0.12 and -0.13 for the judicial and legislative checks conditions, respectively. This implies that the probability that a candidate is selected by a member of the oldest birth cohort decreases by about 12 percentage points, compared to a candidate who holds the pro-democracy view. It appears that younger respondents handed out a lesser punishment for antidemocratic behavior, with AMCEs of around -0.05. 

This is a substantively important difference that implies a much lesser electoral cost handed out by younger constituencies. Note, however, that the estimated AMCEs are negative across the board. In other words, across both antidemocratic attributes and all birth cohorts, there is, on average, an electoral punishment for candidates who violate norms relating to legislative and judicial checks on executive power. Nonetheless, the cohort patterns are evident. In Canada, the results point in a similar direction, but are somewhat less conclusive.

\begin{figure}
  \caption{Approval of a shutdown of the state legislature, by birth cohort and randomized objective of the governor}
  \includegraphics{generation.png}
\end{figure}

The second type of evidence that I bring to bear are results from scenarios that ask respondents about concrete anti-democratic actions from incumbents that they must condone or denounce. By doing away with the more abstract formulations of the survey questions examined above, I hope to elicit more meaningful answers from respondents. I present results from an experiment that asked respondents to evaluate a governor's proposal to shut down the legislature in order to impose his or her preferred policy on coronavirus lockdowns. The experiment randomizes information on the partisanship of the governor (Democrat, Republican, or not mentioned) and the objective of the governor (whether to continue the lockdown or to open up the economy). In all cases, the state legislature was depicted as opposing the governor's plan of action. The respondents were asked to indicate whether they would approve of the governor shutting down the legislature (1 = approve; 0 = disapprove). The results shown in Figure 6 point in a clear direction: there is a roughly linear increase in the probability of supporting the shutdown as we progress from the oldest to the youngest birth cohorts. When the objective of the governor is to continue the lockdown, more than 65% of Americans born in 1990 or later agree that the governor shut down the legislature and impose the continuation of the lockdown. 

## Concluding thoughts

Almost two years ago, I set out to prove Foa and Mounk wrong. Their presentation of World Values Survey data seemed dishonest and a perfect distillation of everything I had been taught not to do. Pouring over different data sources and using a variety of measurement tools, I cannot, in any way, prove them wrong: there is clearly something worth investigation. In particular, the precise mechanisms that underlie the patterns presented in this paper merit further attention. 



\newpage 
# References